Stochastic Wind Power Forecasting

### 2020-05-28 LaTeX_Code/Paper/paper_1line.tex
Latest paper's version

{\color{red} Add AMS Classification.}

@@

\item \textbf{Least Square Minimization:} We consider the observed data $v^{M,N+1}$ with $\Delta v_{j,i}=v_{j,i+1}-v_{j,i}$ and length between observations $\Delta$, where $i\in\{0,\dots,N\}$ and $j\in\{1,\dots,M\}$. $(V_{j,i+1}|v_{j,i})$ is a random variable which conditional mean can be approximated by the solution of the system
\begin{equation*}
\begin{cases}
\dif \E\left[V\right]_{j,i}&=-\theta_t\E\left[V\right]_{j,i}\dif t\\
\E\left[V\right]_{j,i}(t_{j,i})&=v_{j,i},
\end{cases}
\end{equation*}
evaluated in $t_{j,i+1}$, i.e., $\E\left[V\right]_{j,i}(t_{j,i+1})$. Then, the random variable $(V_{j,i+1}-\E\left[V\right]_{j,i}(t_{j,i+1}))$ has zero mean. If we assume that $\theta_t=c\in\R^+$ for all $t\in[t_{j,i},t_{j,i+1}]$, then $\E\left[V(t_{j,i+1})\right]_{j,i}=v_{j,i}e^{-c\Delta}$. If we have a total of $M\times N$ transitions, we can write the regression problem for the conditional mean with $L^2$ loss function as
\begin{equation}
\begin{split}
c^*&\approx\arg\min_{c\ \geq\ 0}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-\E\left[v\right]_{j,i}(t_{j,i+1})\right)^2\right]\\
&=\arg\min_{c\ \geq\ 0}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-v_{j,i}e^{-c\Delta}\right)^2\right]\\
&\approx\arg\min_{c\ \geq\ 0}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-v_{j,i}(1-c\Delta)\right)^2\right].
\end{split}
\label{Eq-1}
\end{equation}
As Equation (\ref{Eq-1}) is convex in $c$ (composition of convex is convex), it is enough to verify the first order optimality conditions. From
\begin{equation*}
\frac{\partial}{\partial c}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-v_{j,i}(1-c\Delta)\right)^2\right]=\sum_{j=1}^M\sum_{i=0}^{N-1}2v_{j,i}\Delta\left(v_{j,i+1}-v_{j,i}(1-c\Delta)\right),
\end{equation*}
follows that
\begin{equation}
c^{*}=\frac{\sum_{j=1}^M\sum_{i=0}^{N-1}v_{j,i}(v_{j,i}-v_{j,i+1})}{\Delta\cdot\sum_{j=1}^M\sum_{i=0}^{N-1}(v_{j,i})^2}.
\label{Eq-4}
\end{equation}

\item \textbf{Quadratic Variation:} We approximate the SDE by its E-M scheme. In particular, we approximate the It\^o quadratic variation with the discrete one:
\begin{itemize}

\item It\^o process quadratic variation: $[V]_t=\int_0^t\sigma_s^2\dif s$.
\item Discrete process quadratic variation: $[V]_t=\sum_{0<s\leq t}(\Delta V_s)^2$.

Then, considering $\Delta$ the time between measurements, we approximate:
\begin{equation}
\theta_0^*\alpha^*\approx\frac{\sum_{j=1}^M\sum_{i=1}^N(\Delta v_{j,i})^2}{2\Delta\cdot\sum_{j=1}^M\sum_{i=1}^N(v_{j,i}+p_{j,i})(1-v_{j,i}-p_{j,i})}.
\label{Eq-2}
\end{equation}

@@

Let $\{\Delta v_{j,i}\}_{j=1,i=1}^{M,N}$ be the set of error observation transitions, and $\psi(\Delta v_{j,i};\bm{\theta})$ the Lamperti transform for the transitions. The transformed observations transitions $\{\Delta z_{j,i}\}_{j=1,i=1}^{M,N}$ depend on the parameter $\bm{\theta}$.\\
The problem of maximizing the  $$\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\Delta z_{j,i}\}_{j=1,i=1}^{M,N}\right),$$
is not totally defined as the data $\{\Delta z_{j,i}\}_{j=1,i=1}^{M,N}$ depend on $\bm{\theta}$. However, we propose to find a fixed point $\bm{\theta}^*$ such that
\begin{equation}
\bm{\theta}^*=\arg\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\psi(\bm{\theta}^*;\Delta v_{j,i})\}_{j=1,i=1}^{M,N}\right).
\label{FP}
\end{equation}
{\color{red}In this point, the likelihood has a maximum for the data set corresponding to that point.}
We are interested in finding solutions for (\ref{FP}). If $\bm{\theta}^*$ solves (\ref{FP}), then
\begin{equation*}
\bm{\theta}^*-\arg\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\psi(\bm{\theta}^*;\Delta v_{j,i})\}_{j=1,i=1}^{M,N}\right)=0.
\end{equation*}
Given $\bm{\theta}^*=(\theta_0^*,\alpha^*)$, we call $\bm{\theta}^{**}=(\theta_0^{**},\alpha^{**})$ to the solution of
\begin{equation*}
\arg\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\psi(\bm{\theta}^*;\Delta v_{j,i})\}_{j=1,i=1}^{M,N}\right).
\end{equation*}
For each $\bm{\theta}^*$, we define the relative error function
\begin{equation}
\Psi(\bm{\theta}^*)=\frac{|\theta_0^{*}-\theta_0^{**}|}{|\theta_0^{*}|}+\frac{|\alpha^{*}-\alpha^{**}|}{|\alpha^{*}|},
\label{eq:rel_error}
\end{equation}
and remark that $\bm{\theta}^*$ is a fixed point if and only if $\Psi(\bm{\theta}^*)=0$.

@@

For any $j=1, \ldots, M$, we extrapolate linearly the truncated prediction function to get the value for $p_{j,-\delta}$, and set $v(t_{j,-\delta})=0$. In this way, we consider the initial transition $(V_{j, 0}|v_{j,-\delta};\bm{\theta},\delta)$. 
We assume that this initial transition has a Beta distribution and apply the same moment matching method as for the rest of the transitions described above. Given an initial guess for $\bm{\theta}$, we obtain the initial estimation for $\delta$ solving the following problem
\begin{equation}
\delta\approx\arg\max_{\delta}\mathcal{L}_{\delta}\left(\bm{\theta},\delta; v^{M,1}\right) = \arg\max_{\delta}\prod\limits_{j=1}^M \rho_0 \left(v_{j, 0}|v_{j,-\delta};\bm{\theta},\delta\right),
\label{eq:likelihood_delta}
\end{equation}
where $\mathcal{L}_\delta$ is the $\delta-$likelihood. To solve this problem, we repeat the steps described in subsection (\ref{Sec:MinLH}), with the additional initial step of creating the linear extrapolation for $p_{j,-\delta}$.\\
\quad\\
As remarked, we can extend the statistical model (\ref{likelihood}) to include the extra parameter $\delta$. The complete likelihood, $\mathcal{L}_c$, is given by
\begin{equation}
\mathcal{L}_c\left(\bm{\theta},\delta; v^{M,N +1}\right)=\mathcal{L}\left(\bm{\theta}; v^{M,N +1}\right)\mathcal{L}_{\delta}\left(\bm{\theta},\delta; v^{M,1}\right).
\label{eq:complete_LH}
\end{equation}
As we may provide initial guesses for $\bm{\theta}$ and $\delta$, we have a starting point for the maximization of (\ref{eq:complete_LH}).

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{../../MATLAB_Files/Results/likelihood/normal/Log-Likelihood.eps}
\caption{Negative log-likelihood's level sets. All optimal values are located over the curve $\theta_0\alpha=0.097$.}
\label{fig:neg-LL}
\end{figure}
On Figure (\ref{fig:neg-LL}), we can see the level sets for the negative log-likelihood. The numerical values of each important point can be seen in Table (\ref{tab:optimal_values}). We set $(\theta_0^E,\alpha^E)=(1.93,0.050)$, as it is where the negative log-likelihood reaches its minimum value.
\begin{table}[H]
\centering
\begin{tabular}{lccc}
\toprule
 & $\theta_0$ & $\alpha$ & $\theta_0\alpha$\\
 \midrule
 Initial guess & 1.54 & 0.070 & 0.111 \\
 fminsearch & 1.14 & 0.076 & 0.097 \\
 fmincon & 1.58 & 0.062 & 0.097 \\
 fminunc & 1.54 & 0.063 & 0.097 \\
 evaluations & 1.93 & 0.050 & 0.097 \\
 \bottomrule
\end{tabular}
\caption{Value points from Figure \ref{fig:neg-LL}.}
\label{tab:optimal_values}
\end{table}
Observe that all the local (possibly global) minimizers are located in curve $\theta_0,\alpha=0.097$. We will explain this effect later, but it is a consequence of Markovianity and the indistinguishable observed transitions.

@@

  \item Model 2: This model features derivative tracking and time-varying mean-reversion parameter:  
%i.e. it is equivalent to (\ref{model:derivative_tracking_X}).
%  \begin{align}
%  dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \alpha (V_t +p_t ) (1-V_t-p_t)} \  dW_t   \nonumber \\ %\quad t > 0
%  V_0 & = v_0 \,, \label{M2}
%  \end{align}
%  with $\theta_t$ given by (\ref{theta_t}).
\begin{equation}
\begin{cases}
\dif X_t&= \big(\dot{p}_t  - \theta_t (X_t - p_t) \big) \dif t +\sqrt{2 \alpha \theta_0 X_t (1 - X_t)} \dif W_t, \quad t \in [0,T]\\
X_0&=x_0\in (0,1) \,,
\end{cases}.  \label{Model2}
\end{equation}
 with $\theta_0 > 0, \, \alpha > 0$ and .
\end{itemize}
To show the better performance of Model 2, we calculate the Akaike information criterion (AIC) and Bayesian information criterion (BIC) for each provider, each model, and different proxies densities. Table (\ref{tab:model_comparison}) summarizes this computations. We can observe that the best fitting is achieved with Model 2 and Beta proxies.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{../../MATLAB_Files/Results/delta/contour_delta.png}
\caption{Initial value for $\delta$ as a function of the elements of the parameter vector $\bm{\theta}$.}
\label{fig:delta}
\end{figure}

>>>>

{\color{red} 2020 AMS Subject Classification: 60H10}

@@

\item \textbf{Least square minimization:} We consider the observed data $v^{M,N+1}$ with $\Delta v_{j,i}=v_{j,i+1}-v_{j,i}$ and length between observations $\Delta$, where $i\in\{0,\dots,N\}$ and $j\in\{1,\dots,M\}$. $(V_{j,i+1}|v_{j,i})$ is a random variable which conditional mean can be approximated by the solution of the system
\begin{equation*}
\begin{cases}
\dif \E\left[V\right]_{j,i}&=-\theta_t\E\left[V\right]_{j,i}\dif t\\
\E\left[V\right]_{j,i}(t_{j,i})&=v_{j,i},
\end{cases}
\end{equation*}
evaluated in $t_{j,i+1}$, i.e., $\E\left[V\right]_{j,i}(t_{j,i+1})$. Then, the random variable $(V_{j,i+1}-\E\left[V\right]_{j,i}(t_{j,i+1}))$ has zero mean. If we assume that $\theta_t=c\in\R^+$ for all $t\in[t_{j,i},t_{j,i+1}]$, then $\E\left[V(t_{j,i+1})\right]_{j,i}=v_{j,i}e^{-c\Delta}$. If we have a total of $M\times N$ transitions, we can write the regression problem for the conditional mean with $L^2$ loss function as
\begin{equation}
\begin{split}
c^*&\approx\arg\min_{c\ \geq\ 0}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-\E\left[v\right]_{j,i}(t_{j,i+1})\right)^2\right]\\
&=\arg\min_{c\ \geq\ 0}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-v_{j,i}e^{-c\Delta}\right)^2\right]\\
&\approx\arg\min_{c\ \geq\ 0}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-v_{j,i}(1-c\Delta)\right)^2\right].
\end{split}
\label{Eq-1}
\end{equation}
As Equation (\ref{Eq-1}) is convex in $c$ (composition of convex is convex), it is enough to verify the first order optimality conditions. From
\begin{equation*}
\frac{\partial}{\partial c}\left[\sum_{j=1}^M\sum_{i=0}^{N-1}\left(v_{j,i+1}-v_{j,i}(1-c\Delta)\right)^2\right]=\sum_{j=1}^M\sum_{i=0}^{N-1}2v_{j,i}\Delta\left(v_{j,i+1}-v_{j,i}(1-c\Delta)\right),
\end{equation*}
it follows that
\begin{equation}
c^{*}=\frac{\sum_{j=1}^M\sum_{i=0}^{N-1}v_{j,i}(v_{j,i}-v_{j,i+1})}{\Delta\cdot\sum_{j=1}^M\sum_{i=0}^{N-1}(v_{j,i})^2}.
\label{Eq-4}
\end{equation}

\item \textbf{Quadratic variation:} We use the numerical Euler-Maruyama (E-M) scheme to approximate the solution to the SDE  \eqref{VtSDE}. In particular, we approximate the quadratic variation of the  It\^{o}'s  process $V$
\begin{equation*}
[V]_t=\int_0^t b(V_s; \bm{\theta}, p_s)^2 \dif s, \text{ where }  \:\: b(V_s; \bm{\theta}, p_s) = 
\sqrt{2 \alpha \theta_0 (V_s +p_s ) (1-V_s - p_s)}\,,
\end{equation*}
with the discrete sum $\sum_{0< t_{j, i-1} \leq t}(V_{t_{j, i}} - V_{t_{j, i-1}})^2$.

As initial guess for the diffusion variability coefficient $\theta_0 \alpha$, we choose
\begin{equation}
\theta_0^*\alpha^*\approx\frac{\sum_{j=1}^M\sum_{i=1}^N(v_{j,i} - v_{j,i-1})^2}{2\Delta\cdot\sum_{j=1}^M\sum_{i=1}^N(v_{j,i}+p_{j,i})(1-v_{j,i}-p_{j,i})}\,,
\label{Eq-2}
\end{equation}
where $\Delta$ is the length of the time interval between two consecutive measurements.
\end{itemize}

@@

Let $\{\Delta v_{j,i}\}_{j=1,i=1}^{M,N}$ be the set of error observation transitions, and $\psi(\Delta v_{j,i};\bm{\theta})$ the Lamperti transform for the transitions. The transformed observations transitions $\{\Delta z_{j,i}\}_{j=1,i=1}^{M,N}$ depend on the parameter $\bm{\theta}$.\\
The problem of maximizing the  $$\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\Delta z_{j,i}\}_{j=1,i=1}^{M,N}\right),$$
is not totally defined as the data $\{\Delta z_{j,i}\}_{j=1,i=1}^{M,N}$ depend on $\bm{\theta}$. However, we propose to find a fixed point $\bm{\theta}^*$ such that
\begin{equation}
\bm{\theta}^*=\arg\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\psi(\bm{\theta}^*;\Delta v_{j,i})\}_{j=1,i=1}^{M,N}\right).
\label{FP}
\end{equation}
{\color{red}In this point, the likelihood has a maximum for the data set corresponding to that point.}
We are interested in finding solutions for (\ref{FP}). If $\bm{\theta}^*$ solves (\ref{FP}), then
\begin{equation*}
\bm{\theta}^*-\arg\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\psi(\bm{\theta}^*;\Delta v_{j,i})\}_{j=1,i=1}^{M,N}\right)=0.
\end{equation*}
Given $\bm{\theta}^*=(\theta_0^*,\alpha^*)$, we call $\bm{\theta}^{**}=(\theta_0^{**},\alpha^{**})$ to the solution of
\begin{equation*}
\arg\max_{\bm{\theta}}\mathcal{L}_L\left(\bm{\theta};\{\psi(\bm{\theta}^*;\Delta v_{j,i})\}_{j=1,i=1}^{M,N}\right).
\end{equation*}
For each $\bm{\theta}^*$, we define the relative error function
\begin{equation}
\Psi(\bm{\theta}^*)=\frac{|\theta_0^*\alpha^*-\theta_0^{**}\alpha^{**}|}{|\theta_0^*\alpha^*|},
\label{eq:rel_error}
\end{equation}
and remark that $\bm{\theta}^*$ is a fixed point if and only if $\Psi(\bm{\theta}^*)=0$.

@@

For any $j=1, \ldots, M$, we extrapolate backward linearly the truncated prediction function to get its value at time $-\delta$,$p_{j,-\delta}$, and set $v_{t_{j,-\delta}}=0$. We assume that the initial transition $(V_{j, 0}|v_{j,-\delta};\bm{\theta},\delta)$ 
has a Beta distribution and apply to it the same moment matching method used above. Given a starting value for $\bm{\theta}$, we obtain the initial estimation for $\delta$ solving the following problem
\begin{equation}
\delta\approx\arg\max_{\delta}\mathcal{L}_{\delta}\left(\bm{\theta},\delta; v^{M,1}\right) = \arg\max_{\delta}\prod\limits_{j=1}^M \rho_0 \left(v_{j, 0}|v_{j,-\delta};\bm{\theta},\delta\right),
\label{eq:likelihood_delta}
\end{equation}
where $\mathcal{L}_\delta$ is the $\delta-$likelihood. To solve this problem, we repeat the steps described in subsection (\ref{Sec:MinLH}), with the additional initial step of creating the linear extrapolation for $p_{j,-\delta}$.\\
\quad\\
As remarked, we extend the statistical model (\ref{likelihood}) to include the extra parameter $\delta$. The complete likelihood, $\mathcal{L}_c$, is given by
\begin{equation}
\mathcal{L}_c\left(\bm{\theta},\delta; v^{M,N +1}\right)=\mathcal{L}\left(\bm{\theta}; v^{M,N +1}\right)\mathcal{L}_{\delta}\left(\bm{\theta},\delta; v^{M,1}\right).
\label{eq:complete_LH}
\end{equation}
As we may provide initial guesses for $\bm{\theta}$ and $\delta$, we have a starting point for the numerical optimization of the complete likelihood (\ref{eq:complete_LH}).

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{../../MATLAB_Files/Results/likelihood/normal/Log-Likelihood.eps}
\caption{Negative log-likelihood's level sets. All optimal values are located over the curve $\theta_0\alpha=0.097$.}
\label{fig:neg-LL}
\end{figure}
On Figure (\ref{fig:neg-LL}), we can see the level sets for the negative log-likelihood. The numerical values of each important point can be seen in Table (\ref{tab:optimal_values}). We set $(\theta_0^E,\alpha^E)=(1.93,0.050)$, as it is where the negative log-likelihood reaches its minimum value.
\begin{table}[H]
\centering
\begin{tabular}{lccc}
\toprule
 & $\theta_0$ & $\alpha$ & $\theta_0\alpha$\\
 \midrule
 Initial guess & 1.54 & 0.070 & 0.111 \\
 fminsearch & 1.14 & 0.076 & 0.097 \\
 fmincon & 1.58 & 0.062 & 0.097 \\
 fminunc & 1.54 & 0.063 & 0.097 \\
 evaluations & 1.93 & 0.050 & 0.097 \\
 \bottomrule
\end{tabular}
\caption{Value points from Figure \ref{fig:neg-LL}.}
\label{tab:optimal_values}
\end{table}
Observe that all the local (possibly global) minimizers are located in curve $\theta_0\alpha=0.097$. We will explain this effect later, but it is a consequence of Markovianity and the indistinguishable observed transitions.

@@

  \item Model 2: This model features derivative tracking and time-varying mean-reversion parameter:  
%i.e. it is equivalent to (\ref{model:derivative_tracking_X}).
%  \begin{align}
%  dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \alpha (V_t +p_t ) (1-V_t-p_t)} \  dW_t   \nonumber \\ %\quad t > 0
%  V_0 & = v_0 \,, \label{M2}
%  \end{align}
%  with $\theta_t$ given by (\ref{theta_t}).
\begin{equation}
\begin{cases}
\dif X_t&= \big(\dot{p}_t  - \theta_t (X_t - p_t) \big) \dif t +\sqrt{2 \alpha \theta_0 X_t (1 - X_t)} \dif W_t, \quad t \in [0,T]\\
X_0&=x_0\in (0,1) \,,
\end{cases}.  \label{Model2}
\end{equation}
 with $\theta_0 > 0, \, \alpha > 0$ and $\theta_t$ satisfying the condition \eqref{Assumption:3} .
\end{itemize}
To show the better performance of Model 2, we calculate the Akaike information criterion (AIC) and Bayesian information criterion (BIC) for each provider, each model, and different proxies densities. Table (\ref{tab:model_comparison}) summarizes this computations. We can observe that the best fitting is achieved with Model 2 and Beta proxies.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{../../MATLAB_Files/Results/delta/contour_delta.eps}
\caption{Initial value for $\delta$ as a function of the elements of the parameter vector $\bm{\theta}$.}
\label{fig:delta}
\end{figure}
